---
title: "Comparisons_against_the_original_R_functions"
author: Suyuan Wang
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparisons_against_the_original_R_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GLM.g.p)
library(haven)
library(bench)
# Load completedata.sas7bdat from the package
data <- read_sas(system.file("data", "completedata.sas7bdat", package = "GLM.g.p"))
```

## Introduction
This vignette demonstrates the usage of the M_gaussian and M_poisson functions in the GLM.g.p package, including comparisons against R's built-in glm() function for Gaussian and Poisson families to validate correctness and efficiency.

## Comparison 1: Gaussian Regression with M_gaussian
We compare the results of M_gaussian() with glm() using the Gaussian family.
```{r}
# Fit models
result_gaussian <- M_gaussian(Depression ~ Fatalism , data = data)
result_glm <- glm(Depression  ~ Fatalism, data = data,family = gaussian)

# Compare coefficients
coef_gaussian <- result_gaussian$coefficients[, "Estimate"]
coef_glm <- coef(result_glm)

cat("Coefficients Comparison:\n")
print(all.equal(coef_gaussian, coef_glm, tolerance = 1e-6))
```
## Efficiency Comparison
We use the bench package to compare the performance of M_gaussian() and glm(family = gaussian).
```{r}
# Benchmarking execution time only
benchmark_gaussian <- bench::mark(
  M_gaussian = {
    M_gaussian(Depression ~ Fatalism, data = data)
    NULL  # Ensure consistent return value
  },
  glm_gaussian = {
    glm(Depression ~ Fatalism, data = data, family = gaussian)
    NULL  # Ensure consistent return value
  },
  iterations = 100
)

# Print benchmark results
print(benchmark_gaussian)

```

## Comparison 2: Poisson Regression with M_poisson
We compare the results of M_poisson() with glm() using the Poisson family.
```{r}
# Fit models
result_poisson <- M_poisson(Comorbidity1  ~ Depression, data = data)
result_glm <- glm(Comorbidity1  ~ Depression, family = poisson, data = data)

# Compare coefficients
coef_poisson <- result_poisson$coefficients
coef_glm <- coef(result_glm)

cat("Coefficients Comparison:\n")
print(all.equal(coef_poisson, coef_glm, tolerance = 1e-6))
```

## Efficiency Comparison
We use the bench package to compare the performance of M_poisson() and glm(family = poisson).

```{r}
# Benchmark comparison for Poisson regression models
benchmark_poisson <- bench::mark(
  M_poisson = {
    M_poisson(Comorbidity1 ~ Fatalism, data = data)
    NULL  # Ensure consistent return value
  },
  glm_poisson = {
    glm(Comorbidity1 ~ Fatalism, family = poisson, data = data)
    NULL  # Ensure consistent return value
  },
  iterations = 100
)

# Print benchmark results
print(benchmark_poisson)
```

## Summary
Both M_gaussian and M_poisson produce results that are nearly identical to the respective built-in glm() functions (family = gaussian and family = poisson).
Benchmarks demonstrate that the custom implementations are comparable in efficiency, with potential trade-offs depending on use cases.

For more details about the functions, refer to their help pages:
```{r}
?M_gaussian
?M_poisson
```



