---
title: "Using GLM.g.p Package"
author: "Suyuan Wang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using GLM.g.p Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r config, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

##Introduction

This vignette demonstrates the usage of the M_gaussian and M_poisson functions from the GLM.g.p package for generalized linear models.
The M_gaussian function implements ordinary least squares regression for Gaussian-distributed data, while the M_poisson function implements Poisson regression for count data using Iteratively Reweighted Least Squares (IRLS).

## Installation
*To install the package, use the following command:*
```{r}
devtools::install_github("/biostat625_hw3/GLM.g.p")
library(GLM.g.p)
```

##Example 1: Gaussian Regression
*The M_gaussian function performs ordinary least squares regression for Gaussian-distributed data. Below is an example.*

Here, we use the completedata.sas7bdat dataset included in the package and model Depression as a function of Fatalism.
```{r}
# Load Example Data
library(haven)
data <- read_sas(system.file("data", "completedata.sas7bdat", package = "GLM.g.p"))
head(data)

# Fit Gaussian regression model
result_gaussian <- M_gaussian(Depression ~ Fatalism, data = data)

# Output results
result_gaussian
```
## Interpretation

The Estimate column contains the regression coefficients.
The R-squared and Adjusted R-squared provide measures of model fit.
The F-statistic and corresponding p-value indicate the overall significance of the model.

## Example 2: Poisson Regression
*The M_poisson function performs Poisson regression, typically used for count data. Below is an example.*

Here, we model Depression as a function of Comorbidity1.
```{r}
# Fit Poisson regression model
result_poisson <- M_poisson(Depression ~ Comorbidity1, data = data)

# Output results
cat("Poisson Regression Results:\n")
cat("Coefficients:\n")
print(result_poisson$coefficients)
cat("\nNull Deviance:", result_poisson$null_deviance, "\n")
cat("Residual Deviance:", result_poisson$residual_deviance, "\n")
cat("Converged:", result_poisson$converged, "\n")
```
## Interpretation

The Coefficients represent the log-linear relationship between the predictor (Comorbidity1) and the response (Depression).
The Null Deviance and Residual Deviance indicate model fit. Lower residual deviance suggests better fit.


# Summary

The `GLM.g.p` package provides two main functions:

1. `M_gaussian`: For linear regression on Gaussian-distributed data.
2. `M_poisson`: For Poisson regression on count data.



